---
title: "Assembly"
output: html_document
date: "2024-06-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# generate trans_nullmodel object
## as an example, we only use high abundance OTU with mean relative abundance > 0.0005
t1 <- trans_nullmodel$new(dataset, filter_thres = 0.0005, add_data = env_data_16S)

## use pH as the test variable
t1$cal_mantel_corr(use_env = "pH")
## return t1$res_mantel_corr
## plot the mantel correlogram
t1$plot_mantel_corr()

## see null.model parameter for other null models
## null model run 500 times for the example
t1$cal_ses_betampd(runs = 500, abundance.weighted = TRUE)
## return t1$res_ses_betampd

## add betaNRI matrix to beta_diversity list
dataset$beta_diversity[["betaNRI"]] <- t1$res_ses_betampd
## create trans_beta class, use measure "betaNRI"
t2 <- trans_beta$new(dataset = dataset, group = "Group", measure = "betaNRI")
## transform the distance for each group
t2$cal_group_distance()
## see the help document for more methods, e.g. "anova" and "KW_dunn"
t2$cal_group_distance_diff(method = "wilcox")
## plot the results
g1 <- t2$plot_group_distance(boxplot_add = "mean")
g1 + geom_hline(yintercept = -2, linetype = 2) + geom_hline(yintercept = 2, linetype = 2)

## null model run 500 times
t1$cal_ses_betamntd(runs = 500, abundance.weighted = TRUE, null.model = "taxa.labels")
## return t1$res_ses_betamntd

tmp <- "./test1"; dir.create(tmp)
t1$cal_ses_betamntd(runs = 1000, abundance.weighted = TRUE, use_iCAMP = TRUE, iCAMP_tempdir = tmp)

## result stored in t1$res_rcbray
t1$cal_rcbray(runs = 1000)
## return t1$res_rcbray

##use betaNTI and rcbray to evaluate processes
t1$cal_process(use_betamntd = TRUE, group = "Group")

t1$cal_process(use_betamntd = TRUE)

t1$res_process

## require NST package to be installed
t1$cal_NST(method = "tNST", group = "Group", dist.method = "bray", abundance.weighted = TRUE, output.rand = TRUE, SES = TRUE)
t1$res_NST$index.grp

## test the NST difference between each pair of groups
t1$cal_NST_test(method = "nst.boot")

##convert long format table to square matrix
## the 10th column: MST.ij.bray in t1$res_NST$index.pair
test <- t1$cal_NST_convert(10)

## for pNST method, phylogenetic tree is needed
t1$cal_NST(method = "pNST", group = "Group", output.rand = TRUE, SES = TRUE)
t1$cal_NST_test(method = "nst.boot")

t1$cal_NRI(null.model = "taxa.labels", abundance.weighted = FALSE, runs = 999)
t1$cal_NTI(null.model = "taxa.labels", abundance.weighted = TRUE, runs = 999)



#ICAMP Package#
##Install/Load Libraries##
```{r}
install.packages(c("permute", "ape", "bigmemory", "nortest", "DirichletReg", "data.table"))

library(vegan)
library(permute)
library(ape)
library(bigmemory)
library(nortest)
library(minpack.lm)
library(Hmisc)
library(stats4)
library(DirichletReg)
library(data.table)
library(ggVennDiagram)
library(VennDiagram)
library(ggpubr)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(phyloseq)
library(microbiome)
library(ggrepel)

##Install Color Palettes
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
ibm.cbb <- c("#648FFF", "#785EF0", "#DC267F", "#FE6100", "grey")
tol.cbb <- c("#332288", "#117733", "#44AA99", "#88CCEE", "#DDCC77", "#CC6677", "#AA4499", "#882255")

```






