---
title: "Diff_Abundance"
output: html_document
date: "2024-06-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##### Differential Abundance #####
```{r}
###### Libraries #####
library(phyloseq)
library(microbiome)
library(vegan)
library(tidyverse)
library(ggplot2)
library(minpack.lm)
library(Hmisc)
library(stats4)
library(ggrepel)
library(ANCOMBC)
library(ggVennDiagram)
library(VennDiagram)
library(ggpubr)
library(dplyr)

# set options for scientific numbers to not be displayed
options(scipen=10000) 

# color blind pallet used throughout 
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
ibm.cbb <- c("#648FFF", "#785EF0", "#DC267F", "#FE6100", "grey")
tol.cbb <- c("#332288", "#117733", "#44AA99", "#88CCEE", "#DDCC77", "#CC6677", "#AA4499", "#882255")
```
#### Read in RDS file #### 
```{r}
# using the non-normalized reads since spieceasi has its own normalizaiton methods

bac_sperm <- readRDS("2024-01-24_SpermosphereSteaming_Bacteria/RDS/Bacteria_spermosphere_nonnorm_2024-01-24.rds") ##Using Non-normalized Reads
```
#### diff abundance with ancom-BC2 - don't need to run. load the RDS file ####

###UNSURE LEAVING THIS OUT OF CHUNKS FOR NOW### #I am doing diff abundance testing with 12 and 18 hours since there was no effect of time between these two time points. 
#Steamed Subsetting
# In other words 12 and 18 were similar composition after 12 hours. 



##STEAMED##
```{r}
b <- bac_sperm %>%
  psmelt() %>% #combining the Physeq Object
  group_by(OTU, Trt, Time, Type) %>% #Grouping by different 
  dplyr::summarize(sumabund = sum(Abundance)) #Has Zeros in It? (dpylr :: makes it use the command from that)

b$PA <- ifelse(b$sumabund > 0, 1, 0)
Soybean_Epiphyte.present <- b$OTU[b$PA == 1 & b$Trt == "Soybean Epiphytes"]
BulkSoil_Planting_Steamed.present <- b$OTU[b$PA == 1 & b$Trt == "Steamed " & b$Time == "Planting" & b$Type == "Bulk.Soil"]
b1 <- intersect(Soybean_Epiphyte.present, BulkSoil_Planting_Steamed.present)

c <- bac_sperm %>%
      subset_taxa(OTU %in% b1) %>%
      subset_samples(Type %in% c("Spermosphere", "Rhizosphere") & Trt %in% "Steamed ") %>%
      phyloseq::filter_taxa(function(x) sum(x) > 0, TRUE)
##SPACE AFTER STEAMED IN METADATA
#### taxonomy table ####
tax.bac <- c %>%
  tax_table() %>%
  as.data.frame()

```


### takes a while to finish -  maybe 30 min
```{r}
out = ancombc2(data = c, 
              assay_name = NULL,
              p_adj_method = "holm", 
              prv_cut = 0.50, 
              fix_formula = "Time",
              group = "Time", 
              struc_zero = TRUE, 
              neg_lb = TRUE, 
              alpha = 0.05, 
              global = TRUE, 
              n_cl = 1, verbose = TRUE)

saveRDS(out, "SteamedDiffAbund.rds")
#Q is the adjusted p-value from holm method

```
### diff abundance test ###
```{r}

diff.abund <- readRDS("SteamedDiffAbund2.rds") ##Using Non-normalized Reads

diff.abund <- out$res
diff.abund2 <- left_join(diff.abund, tax.bac, by = c("taxon" = "OTU"))
diff.abund2$diff_abund <- interaction(diff.abund2$diff_Time17hrs, diff.abund$diff_TimeV2)
diff.abund2$Crop_diff_abund <- ifelse(diff.abund2$diff_abund == "TRUE.TRUE", "Difference Across Both Times", 
                                      ifelse(diff.abund2$diff_abund == "TRUE.FALSE", "Spermosphere Only",
                                             ifelse(diff.abund2$diff_abund == "FALSE, FALSE", "No Difference", 
                                                    ifelse(diff.abund2$diff_abund == "FALSE, TRUE", "Rhizosphere Only", "Not Difference"))))

##BELOW IS OLD CODE
diff.abund2$Crop_diff_abund <- ifelse(diff.abund2$diff_abund == "TRUE.TRUE.TRUE", "Difference Across All Times", 
                                      ifelse(diff.abund2$diff_abund == "TRUE.TRUE.FALSE", "Spermosphere Only",
                                             ifelse(diff.abund2$diff_abund == "TRUE.FALSE.FALSE", "Planting Only",
                                                    ifelse(diff.abund2$diff_abund == "FALSE.FALSE.FALSE", "No Difference Across All Times",
                                                           ifelse(diff.abund2$diff_abund == "FALSE.FALSE.TRUE", "Rhizosphere Only",
                                                                  ifelse(diff.abund2$diff_abund == "FALSE.TRUE.FALSE", "17 Hours Only",
                                                                         ifelse(diff.abund2$diff_abund == "FALSE.TRUE.TRUE", "17 Hours and Rhizosphere Only",
                                                                                ifelse(diff.abund2$diff_abund == "TRUE.FALSE.TRUE", "Planting and Rhizosphere Only", "Not Different"))))))))

saveRDS(diff.abund2, "SteamedDiffAbund2.rds")
diff.abund2 <- readRDS("SteamedDiffAbund2.rds")
```
### supplemental table 2
```{r}
write.csv(diff.abund2, "diff_abund.csv")

diff.abund2 %>%
  subset(Crop_diff_abund != "No Difference") %>%
  group_by(Phylum) %>%
  summarise(n = n()) %>%
  mutate(freq = round(n / sum(n), 2))
  
diff.abund2 %>%
  subset(Crop_diff_abund != "No Difference" & Phylum == "Firmicutes") %>%
  group_by(Genus) %>%
  summarise(n = n()) %>%
  mutate(freq = round(n / sum(n), 2)) %>%
  arrange(-freq)

diff.abund2 %>%
  subset(Crop_diff_abund != "No Difference" & Phylum == "Proteobacteria") %>%
  group_by(Genus) %>%
  summarise(n = n()) %>%
  mutate(freq = round(n / sum(n), 2)) %>%
  arrange(-freq)

diff.abund2 %>%
  subset(Crop_diff_abund != "No Difference") %>%
  group_by(Crop_diff_abund) %>%
  summarise(n = n()) %>%
  mutate(freq = round(n / sum(n), 2)) %>%
  arrange(-freq)

diff.abund2 %>%
  subset(Crop_diff_abund != "No Difference" & Phylum == "Proteobacteria") %>%
  group_by(Genus) %>%
  summarise(n = n()) %>%
  mutate(freq = round(n / sum(n), 2)) %>%
  arrange(-freq) %>%
  print(n = 50)


diff.aubndant.composition <- diff.abund2 %>%
  subset(Crop_diff_abund != "No Difference") %>%
ggplot(aes(x = Crop_diff_abund, fill = Genus)) +
  geom_bar() +
  scale_fill_manual(values = c(cbbPalette, ibm.cbb, tol.cbb)) +
  theme_classic() + 
  xlab("")+
  ylab("Count") +
  theme(axis.text.x = element_text(angle=45, hjust=1),
        legend.text = element_text(face = "italic", size = 5),
        legend.title = element_blank(),
        legend.key.size = unit(0.3, 'cm')) 


# TRUE.TRUE = cotton and soybean 
# TRUE.FALSE = cotton only 
# FALSE.TRUE = soybean only
```


###PLOT FROM ABOVE###
```{r}
diff.abund.plot <- ggplot() + 
  geom_point(data = subset(diff.abund2, diff_abund == "FALSE.FALSE"), aes(x = lfc_CropSoybean, y = `lfc_CropCotton `), color = "grey", shape = 21) +
  geom_point(data = subset(diff.abund2,  diff_abund == "TRUE.FALSE"), aes(x = lfc_CropSoybean, y = `lfc_CropCotton `, fill = Order, size = -log10(`q_CropCotton `)), shape = 22, alpha = 0.7) +
  geom_point(data = subset(diff.abund2,  diff_abund == "FALSE.TRUE"), aes(x = lfc_CropSoybean, y = `lfc_CropCotton `, fill = Order, size = -log10(q_CropSoybean)), shape = 23, alpha = 0.7) +
  geom_point(data = subset(diff.abund2,  diff_abund == "TRUE.TRUE."), aes(x = lfc_CropSoybean, y = `lfc_CropCotton `, fill = Order, size = -log10(q_CropSoybean)), shape = 24, alpha = 0.7) +
  theme_classic() + 
  geom_hline(yintercept = 0, lty = "dotted") +
  geom_vline(xintercept = 0, lty = "dotted") + 
  scale_fill_manual(values = c(cbbPalette, ibm.cbb, tol.cbb)) + 
  xlab("log fold change soybean - soil") + 
  ylab("log fold change cotton - soil") +
  guides(size = FALSE) + 
  theme(legend.position = "right")
  #geom_text_repel(data = subset(diff.abund2,  diff_abund == "TRUE.TRUE"), 
                  #aes(label = Label, x = lfc_CropSoybean, y = `lfc_CropCotton `, fill = Phylum), size = 3)
ggarrange(diff.abund.plot, diff.aubndant.composition)
```


##NONSTEAMED##
```{r}
b$PA <- ifelse(b$sumabund > 0, 1, 0)
Soybean_Epiphyte.present <- b$OTU[b$PA == 1 & b$Trt == "Soybean Epiphytes"]
BulkSoil_Planting_NonSteamed.present <- b$OTU[b$PA == 1 & b$Trt == "Non.Steamed" & b$Time == "Planting" & b$Type == "Bulk.Soil"]
nb1 <- intersect(Soybean_Epiphyte.present, BulkSoil_Planting_NonSteamed.present)

nc <- bac_sperm %>%
      subset_taxa(OTU %in% nb1) %>%
      subset_samples(Type %in% c("Spermosphere", "Rhizosphere") & Trt %in% "Non.Steamed") %>%
      phyloseq::filter_taxa(function(x) sum(x) > 0, TRUE)
##SPACE AFTER STEAMED IN METADATA
#### taxonomy table ####
ntax.bac <- nc %>%
  tax_table() %>%
  as.data.frame()
```
# takes a while to finish -  maybe 30 min
```{r}
out2 = ancombc2(data = nc, 
              assay_name = NULL,
              p_adj_method = "holm", 
              prv_cut = 0.50, 
              fix_formula = "Time",
              group = "Time", 
              struc_zero = TRUE, 
              neg_lb = TRUE, 
              alpha = 0.05, 
              global = TRUE, 
              n_cl = 1, verbose = TRUE)

saveRDS(out2, "NonSteamedDiffAbund.rds")
#Q is the adjusted p-value from holm method

```
###### diff abundance test #####
```{r}
ndiff.abund <- out2$res
ndiff.abund2 <- left_join(ndiff.abund, ntax.bac, by = c("taxon" = "OTU"))
ndiff.abund2$diff_abund <- interaction(ndiff.abund2$diff_Time17hrs, ndiff.abund$diff_TimeV2)
ndiff.abund2$Crop_diff_abund <- ifelse(ndiff.abund2$diff_abund == "TRUE.TRUE", "Difference Across Both Times", 
                                      ifelse(ndiff.abund2$diff_abund == "TRUE.FALSE", "Spermosphere Only",
                                             ifelse(ndiff.abund2$diff_abund == "FALSE, FALSE", "No Difference", 
                                                    ifelse(ndiff.abund2$diff_abund == "FALSE, TRUE", "Rhizosphere Only", "Not Difference"))))

saveRDS(ndiff.abund2, "NonSteamedDiffAbund2.rds")
ndiff.abund2 <- readRDS("NonSteamedDiffAbund2.rds")
```
# supplemental table 2
```{r}
write.csv(diff.abund2, "ndiff_abund.csv")

ndiff.abund2 %>%
  subset(Crop_diff_abund != "No Difference") %>%
  group_by(Phylum) %>%
  summarise(n = n()) %>%
  mutate(freq = round(n / sum(n), 2))
  
ndiff.abund2 %>%
  subset(Crop_diff_abund != "No Difference" & Phylum == "Firmicutes") %>%
  group_by(Genus) %>%
  summarise(n = n()) %>%
  mutate(freq = round(n / sum(n), 2)) %>%
  arrange(-freq)

ndiff.abund2 %>%
  subset(Crop_diff_abund != "No Difference" & Phylum == "Proteobacteria") %>%
  group_by(Genus) %>%
  summarise(n = n()) %>%
  mutate(freq = round(n / sum(n), 2)) %>%
  arrange(-freq)

ndiff.abund2 %>%
  subset(Crop_diff_abund != "No Difference") %>%
  group_by(Crop_diff_abund) %>%
  summarise(n = n()) %>%
  mutate(freq = round(n / sum(n), 2)) %>%
  arrange(-freq)

ndiff.abund2 %>%
  subset(Crop_diff_abund != "No Difference" & Phylum == "Proteobacteria") %>%
  group_by(Genus) %>%
  summarise(n = n()) %>%
  mutate(freq = round(n / sum(n), 2)) %>%
  arrange(-freq) %>%
  print(n = 50)


ndiff.aubndant.composition <- ndiff.abund2 %>%
  subset(Crop_diff_abund != "No Difference") %>%
ggplot(aes(x = Crop_diff_abund, fill = Genus)) +
  geom_bar() +
  scale_fill_manual(values = c(cbbPalette, ibm.cbb, tol.cbb)) +
  theme_classic() + 
  xlab("")+
  ylab("Count") +
  theme(axis.text.x = element_text(angle=45, hjust=1),
        legend.text = element_text(face = "italic", size = 5),
        legend.title = element_blank(),
        legend.key.size = unit(0.3, 'cm')) 


# TRUE.TRUE = cotton and soybean 
# TRUE.FALSE = cotton only 
# FALSE.TRUE = soybean only
```


###PLOT FROM ABOVE###
```{r}
ndiff.abund.plot <- ggplot() + 
  geom_point(data = subset(ndiff.abund2, diff_abund == "FALSE.FALSE"), aes(x = lfc_CropSoybean, y = `lfc_CropCotton `), color = "grey", shape = 21) +
  geom_point(data = subset(ndiff.abund2,  diff_abund == "TRUE.FALSE"), aes(x = lfc_CropSoybean, y = `lfc_CropCotton `, fill = Order, size = -log10(`q_CropCotton `)), shape = 22, alpha = 0.7) +
  geom_point(data = subset(ndiff.abund2,  diff_abund == "FALSE.TRUE"), aes(x = lfc_CropSoybean, y = `lfc_CropCotton `, fill = Order, size = -log10(q_CropSoybean)), shape = 23, alpha = 0.7) +
  geom_point(data = subset(ndiff.abund2,  diff_abund == "TRUE.TRUE."), aes(x = lfc_CropSoybean, y = `lfc_CropCotton `, fill = Order, size = -log10(q_CropSoybean)), shape = 24, alpha = 0.7) +
  theme_classic() + 
  geom_hline(yintercept = 0, lty = "dotted") +
  geom_vline(xintercept = 0, lty = "dotted") + 
  scale_fill_manual(values = c(cbbPalette, ibm.cbb, tol.cbb)) + 
  xlab("log fold change soybean - soil") + 
  ylab("log fold change cotton - soil") +
  guides(size = FALSE) + 
  theme(legend.position = "right")
  #geom_text_repel(data = subset(diff.abund2,  diff_abund == "TRUE.TRUE"), 
                  #aes(label = Label, x = lfc_CropSoybean, y = `lfc_CropCotton `, fill = Phylum), size = 3)
ggarrange(ndiff.abund.plot, ndiff.aubndant.composition)
```







