---
title: "Networking Analysis"
output: html_document
date: "2024-06-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

######Network#########
```{r}
###### Libraries #####
library(igraph)
library(phyloseq)
library(microbiome)
library(vegan)
library(tidyverse)
library(ggplot2)
library(SpiecEasi)
library(NetCoMi)

#### Other packages needed, but installed on the HPC #### 
# NetCoMi, SPRING

# set options for scientific numbers to not be displayed
options(scipen=10000) 

# color blind pallet used throughout 
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
ibm.cbb <- c("#648FFF", "#785EF0", "#DC267F", "#FE6100", "#FFB000")
tol.cbb <- c("#332288", "#117733", "#44AA99", "#88CCEE", "#DDCC77", "#CC6677", "#AA4499", "#882255")
```
#### Read in RDS file #### 
# using the non-normalized reads since spieceasi has its own normalizaiton methods
bac_sperm <- readRDS(file = "Bacteria/Bacteria_spermosphere_nonnorm_112922.rds")
fungi_sperm <- readRDS(file = "Fungi/Fungi_spermosphere_unedited_083022.rds")

# differentially abundant bacteria 
diff.abund <- readRDS("Differential_Abundance/differential_abund_121422.rds")

##### Filter the datasets to what is in the netcomi.R script for taxa assignment/lookup ####

# have to remove the samples that are not the same between the two datasets
bac_sperm_composition_filt_taxa <- bac_sperm %>%  
  phyloseq::subset_samples(Crop %in% c("Cotton ", "Soybean", "Bulk Soil") & !Fungal_Code %in% c("SB.6.5", "C.12.6") & Time.Point %in% c("12", "18")) %>%
  microbiome::transform("compositional") %>%
  microbiome::core(detection = 0.001, prevalence = 0.9) %>%
  taxa_names()

network.bacteria <- bac_sperm %>%
  phyloseq::subset_samples(Crop %in% c("Cotton ", "Soybean", "Bulk Soil") & !Fungal_Code %in% c("SB.6.5", "C.12.6") & Time.Point %in% c("12", "18")) %>%
  subset_taxa(OTU %in% bac_sperm_composition_filt_taxa)

fungi_sperm_composition_filt_taxa <- fungi_sperm %>%  
  subset_samples(Crop %in% c("Cotton ", "Soybean", "Bulk Soil") & !Code %in% c("S.18.5", "C.18.6") & Time.Point %in% c("12", "18")) %>%
  microbiome::transform("compositional") %>%
  microbiome::core(detection = 0.001, prevalence = 0.9) %>%
  taxa_names()

network.fungi <- fungi_sperm %>% 
  subset_samples(Crop %in% c("Cotton ", "Soybean", "Bulk Soil") & !Code %in% c("S.18.5", "C.18.6") & Time.Point %in% c("12", "18")) %>%
  subset_taxa(OTU %in% fungi_sperm_composition_filt_taxa)

# Store count matrices (taxa are columns)
counts_network.bacteria <- as.matrix(t(phyloseq::otu_table(network.bacteria)@.Data))
counts_network.fungi <- as.matrix(t(phyloseq::otu_table(network.fungi)@.Data))

# setting the rownames of the count matrices to the same codes between the two sample sources for cross-domain interactions in spieceasi
rownames(counts_network.bacteria) <- network.bacteria@sam_data$Fungal_Code

# Sanity check - has to be true to proceed. 
all.equal(rownames(counts_network.bacteria), rownames(counts_network.fungi))

# Subsetting based on cotton versus soybean

##### RUN ON HPC #####
# Bacteria
rows_keep_cotton_bacteria <- grep("C", network.bacteria@sam_data$Fungal_Code) 
data_cotton_bacteria <- counts_network.bacteria[rows_keep_cotton_bacteria, ]  # Extract rows from data

rows_keep_soybean_bacteria <- grep("SB", network.bacteria@sam_data$Fungal_Code) 
data_soybean_bacteria <- counts_network.bacteria[rows_keep_soybean_bacteria, ]  # Extract rows from data

rows_keep_soil_bacteria <- grep("S.", network.bacteria@sam_data$Fungal_Code, fixed = T) 
data_soil_bacteria <- counts_network.bacteria[rows_keep_soil_bacteria, ]  # Extract rows from data

# Fungi
rows_keep_cotton_fungi <- grep("C", network.fungi@sam_data$Code) 
data_cotton_fungi <- counts_network.fungi[rows_keep_cotton_fungi, ]  # Extract rows from data

rows_keep_soybean_fungi <- grep("SB", network.fungi@sam_data$Code) 
data_soybean_fungi <- counts_network.fungi[rows_keep_soybean_fungi, ]  # Extract rows from data

rows_keep_soil_fungi <- grep("S.", network.fungi@sam_data$Code, fixed = T) 
data_soil_fungi <- counts_network.fungi[rows_keep_soil_fungi, ]  # Extract rows from data

set.seed(123456)
# Run SpiecEasi and create association matrix for group 1
spiec_result_gr1 <- multi.spiec.easi(list(data_cotton_bacteria, data_cotton_fungi), 
                                     method='mb', nlambda=100, 
                                     lambda.min.ratio=1e-1, 
                                     pulsar.params = list(rep.num = 100))
getStability(spiec_result_gr1) # 0.04918979 - good
saveRDS(spiec_result_gr1, "Networks/cotton_spieceasi_network.rds")

assoMat1 <- SpiecEasi::symBeta(SpiecEasi::getOptBeta(spiec_result_gr1), mode = "ave")

# Run SpiecEasi and create association matrix for group 2
spiec_result_gr2 <- multi.spiec.easi(list(data_soybean_bacteria, data_soybean_fungi), 
                                     method='mb', nlambda=100, 
                                     lambda.min.ratio=1e-1, 
                                     pulsar.params = list(rep.num = 100))
getStability(spiec_result_gr2) # 0.04911035 - good
saveRDS(spiec_result_gr2, "Networks/soybean_spieceasi_network.rds")

assoMat2 <- SpiecEasi::symBeta(SpiecEasi::getOptBeta(spiec_result_gr2), mode = "ave")

# Run SpiecEasi and create association matrix for group 3
spiec_result_gr3 <- multi.spiec.easi(list(data_soil_bacteria, data_soil_fungi), 
                                     method='mb', nlambda=100, 
                                     lambda.min.ratio=1e-1, 
                                     pulsar.params = list(rep.num = 100))
getStability(spiec_result_gr3) # 0.04918979 - good
saveRDS(spiec_result_gr3, "Networks/soil_spieceasi_network.rds")

assoMat3 <- SpiecEasi::symBeta(SpiecEasi::getOptBeta(spiec_result_gr3), mode = "ave")

taxnames <- c(taxa_names(network.bacteria), taxa_names(network.fungi))

#cotton
assoMat1 <- as.matrix(assoMat1)
colnames(assoMat1) <- rownames(assoMat1) <- taxnames
diag(assoMat1) <- 1

#soybean
colnames(assoMat2) <- rownames(assoMat2) <- taxnames
assoMat2 <- as.matrix(assoMat2)
diag(assoMat2) <- 1

#soil
colnames(assoMat3) <- rownames(assoMat3) <- taxnames
assoMat3 <- as.matrix(assoMat3)
diag(assoMat3) <- 1


#### Cotton versus Soybean #####
# Network construction (pass association matrices to netConstruct)
# - sparsMethod must be set to "none" because sparsification is already included in SpiecEasi
sperm.crossdomain.cottonvsoybean <- netConstruct(data = assoMat1, data2 = assoMat2, 
                                                 dataType = "condDependence", 
                                                 sparsMethod = "none")

# Network analysis
netprops_sperm.crossdomain.cottonvsoybean <- netAnalyze(sperm.crossdomain.cottonvsoybean, hubPar = "eigenvector")

# Network comparison
# - Permutation tests cannot be performed because the association matrices are
#   used for network construction. For permutation tests, however, the count 
#   data are needed.
netcomp_soybean_cotton <- netCompare(netprops_sperm.crossdomain.cottonvsoybean, permTest = FALSE, seed = 123456)

##### Cotton versus soil #####
sperm.crossdomain.cottonvsoil <- netConstruct(data = assoMat1, data2 = assoMat3, 
                                                 dataType = "condDependence", 
                                                 sparsMethod = "none")

# Network analysis
netprops_sperm.crossdomain.cottonvsoil <- netAnalyze(sperm.crossdomain.cottonvsoil, hubPar = "eigenvector")

# Network comparison
# - Permutation tests cannot be performed because the association matrices are
#   used for network construction. For permutation tests, however, the count 
#   data are needed.
netcomp_cotton_soil <- netCompare(netprops_sperm.crossdomain.cottonvsoil, permTest = FALSE, seed = 123456)

##### Soybean versus soil #####
sperm.crossdomain.soybeanvssoil <- netConstruct(data = assoMat2, data2 = assoMat3, 
                                                 dataType = "condDependence", 
                                                 sparsMethod = "none")

# Network analysis
netprops_sperm.crossdomain.soybeanvssoil <- netAnalyze(sperm.crossdomain.soybeanvssoil, hubPar = "eigenvector")

# Network comparison
# - Permutation tests cannot be performed because the association matrices are
#   used for network construction. For permutation tests, however, the count 
#   data are needed.
netcomp_soybean_soil <- netCompare(netprops_sperm.crossdomain.soybeanvssoil, permTest = FALSE, seed = 123456)

#### Summary outputs ####
summary(netcomp_soybean_cotton, groupNames = c("Cotton", "Soybean"))
summary(netcomp_cotton_soil, groupNames = c("Cotton", "Soil"))
summary(netcomp_soybean_soil, groupNames = c("Soybean", "Soil"))

saveRDS(sperm.crossdomain.cottonvsoybean, "Networks/netConstruct_cottonvssoybean.rds")
saveRDS(netprops_sperm.crossdomain.cottonvsoybean, "Networks/netanalyse_cottonvssoybean.rds")
saveRDS(netcomp_soybean_cotton, "Networks/netCompare_cottonvssoybean.rds")

saveRDS(sperm.crossdomain.cottonvsoil, "Networks/netConstruct_cottonvsoil.rds")
saveRDS(netprops_sperm.crossdomain.cottonvsoil, "Networks/netanalyse_cottonvsoil.rds")
saveRDS(netcomp_cotton_soil, "Networks/netCompare_cotton_soil.rds")

saveRDS(sperm.crossdomain.soybeanvssoil, "Networks/netConstruct_soybeanvssoil.rds")
saveRDS(netprops_sperm.crossdomain.soybeanvssoil, "Networks/netanalyse_soybeanvssoil.rds")
saveRDS(netcomp_soybean_soil, "Networks/netCompare_soybean_soil.rds")

##### Generate network plots ####
nodeCols <- c(rep(cbbPalette[[3]], ntaxa(network.bacteria)), rep(cbbPalette[[7]], ntaxa(network.fungi)))
nodeShape1 <- c(rep("circle", ntaxa(network.bacteria)), rep("triangle", ntaxa(network.fungi)))

nodeShapeClass <- c(rep("circle", 8), rep("triangle", 8), rep("square", 8), rep("diamond", 8))
labels.label <- as.vector(c(tax_table(network.bacteria)[, "Label"], tax_table(network.fungi)[, "Label"]))
labels.class <- as.vector(c(tax_table(network.bacteria)[, "Class"], tax_table(network.fungi)[, "Class"]))
class.colors <- c(cbbPalette, cbbPalette, cbbPalette, cbbPalette)
color.classes <- class.colors[match(labels.class, unique(labels.class))]
shape.classes <- nodeShapeClass[match(labels.class, unique(labels.class))]

# differential abundant taxa that were included in the networks
diff.abund.taxa.all <- diff.abund$taxon[diff.abund$diff_abund %in% c("TRUE.TRUE", "TRUE.FALSE", "FALSE.TRUE")]
diff.abund.networks <- intersect(bac_sperm_composition_filt_taxa, diff.abund.taxa.all)
diff.abund.cotton <- diff.abund$taxon[diff.abund$diff_abund %in% c("TRUE.TRUE", "TRUE.FALSE")]
diff.abund.soybean <- diff.abund$taxon[diff.abund$diff_abund %in% c("TRUE.TRUE", "FALSE.TRUE")]

labels.order <- as.vector(c(tax_table(network.bacteria)[, "Order"], tax_table(network.fungi)[, "Order"]))
unique(labels.order)

taxnames <- c(taxa_names(network.bacteria), taxa_names(network.fungi))
names(nodeShape1) <- taxnames

names(nodeCols) <- taxnames

names(labels.label) <- taxnames

names(labels.class) <- taxnames

names(color.classes) <- taxnames

names(shape.classes) <- taxnames
# Relative abundance for fungi
otu_rel_fungi_soybean <- as.matrix(as.data.frame(decostand(data_soybean_fungi, method="total", MARGIN=2)))     # mean relative abundance
otu_rel_fungi_cotton <- as.matrix(as.data.frame(decostand(data_cotton_fungi, method="total", MARGIN=2)))     # mean relative abundance

# Relative abundance for bacteria
otu_rel_bac_soybean <- as.matrix(as.data.frame(decostand(data_soybean_bacteria, method="total", MARGIN=2)))
otu_rel_bac_cotton <- as.matrix(as.data.frame(decostand(data_cotton_bacteria, method="total", MARGIN=2)))     # mean relative abundance

# Setting the count matricies equal to relative abundance for cotton or soybean for plotting
netprops_sperm.crossdomain.cottonvsoybean$input$countMat1 <- cbind(100*otu_rel_bac_cotton, 100*otu_rel_fungi_cotton) 
netprops_sperm.crossdomain.cottonvsoybean$input$countMat2 <- cbind(100*otu_rel_bac_soybean, 100*otu_rel_fungi_soybean) 

# Just for manual labeling purposes to figure out where OTUs are. 
plot(netprops_sperm.crossdomain.cottonvsoybean, 
     cexLabels = 5)

#### Figure 6 ####

# cotton versus soybean visual
set.seed(12356)
plot1 <- plot(netprops_sperm.crossdomain.cottonvsoybean, 
     nodeColor = "colorVec",
     colorVec = color.classes,
     nodeSize = "eigenvector",
     nodeSizeSpread = 1,
     nodeFilter = "clustMin",
     nodeFilterPar = 5,
     labels = NULL,
     posCol = cbbPalette[[1]],
     negCol = cbbPalette[[2]],
     sameLayout = F,
     layout = "spring",
     groupNames = c("", ""),
     nodeShape = c("circle", "diamond", "square", "triangle"),
     featVecShape = shape.classes,
     cexLabels = 5,
     cexHubLabels = 5,
     hubTransp = 10,
     mar = c(10,2,2,2))
legend(-2, -1.5, cex = 0.9, pt.cex = 1, 
       legend = unique(labels.class), 
       col = class.colors, 
       bty = "n", pch = c(rep(16, 8),rep(17, 8), rep(15, 8), rep(18, 8)), horiz = F,
       ncol = 3, x.intersp =0.8, y.intersp = 0.8, 
       text.width=c(0.9,0.8,0,1)) 

## different view to get the less connected components which I will add manually to the plot
plot1.2 <- plot(netprops_sperm.crossdomain.cottonvsoybean, 
              nodeColor = "colorVec",
              colorVec = color.classes,
              nodeSize = "eigenvector",
              nodeSizeSpread = 1,
              nodeFilter = "clustMin",
              nodeFilterPar = 1,
              labels = NULL,
              posCol = cbbPalette[[1]],
              negCol = cbbPalette[[2]],
              sameLayout = F,
              layout = "layout_with_kk",
              groupNames = c("", ""),
              nodeShape = c("circle", "diamond", "square", "triangle"),
              featVecShape = shape.classes,
              cexLabels = 0,
              cexHubLabels = 0,
              hubTransp = 10,
              mar = c(10,2,2,2))
legend(-2, -1.5, cex = 0.9, pt.cex = 1, 
       legend = unique(labels.class), 
       col = class.colors, 
       bty = "n", pch = c(rep(16, 8),rep(17, 8), rep(15, 8), rep(18, 8)), horiz = F,
       ncol = 3, x.intersp =0.8, y.intersp = 0.8, 
       text.width=c(0.9,0.8,0,1)) 

# cotton versus soil visual
set.seed(123456)
plot2 <- plot(netprops_sperm.crossdomain.cottonvsoil, 
              nodeColor = "colorVec",
              colorVec = color.classes,
              nodeSize = "eigenvector",
              nodeSizeSpread = 1,
              nodeFilter = "clustMin",
              nodeFilterPar = 5,
              labels = NULL,
              posCol = cbbPalette[[1]],
              negCol = cbbPalette[[2]],
              sameLayout = F,
              layout = "spring",
              groupNames = c("", ""),
              nodeShape = c("circle", "diamond", "square", "triangle"),
              featVecShape = shape.classes,
              cexLabels = 0,
              cexHubLabels = 0,
              hubTransp = 10,
              mar = c(10,2,2,2))
legend(-2, -1.2, cex = 0.9, pt.cex = 1, 
       legend = unique(labels.class), 
       col = class.colors, 
       bty = "n", pch = c(rep(16, 8),rep(15, 8), rep(17, 8)), horiz = F,
       ncol = 3, x.intersp =0.1, y.intersp = 0.5, 
       text.width=c(0,1,0,0.2)) 

plot2.1 <- plot(netprops_sperm.crossdomain.cottonvsoil, 
              nodeColor = "colorVec",
              colorVec = color.classes,
              nodeSize = "eigenvector",
              nodeSizeSpread = 1,
              nodeFilter = "clustMin",
              nodeFilterPar = 1,
              labels = NULL,
              posCol = cbbPalette[[1]],
              negCol = cbbPalette[[2]],
              sameLayout = F,
              layout = "layout_with_kk",
              groupNames = c("", ""),
              nodeShape = c("circle", "diamond", "square", "triangle"),
              featVecShape = shape.classes,
              cexLabels = 0,
              cexHubLabels = 0,
              hubTransp = 10,
              mar = c(10,2,2,2))
legend(-2, -1.2, cex = 0.9, pt.cex = 1, 
       legend = unique(labels.class), 
       col = class.colors, 
       bty = "n", pch = c(rep(16, 8),rep(15, 8), rep(17, 8)), horiz = F,
       ncol = 3, x.intersp =0.1, y.intersp = 0.5, 
       text.width=c(0,1,0,0.2)) 

##### Hub taxa ####
tax.bac <- data.frame(network.bacteria@tax_table)
tax.fun <- data.frame(network.fungi@tax_table)

# Bacteria
soybean.bac.hubs <- tax.bac %>%
  subset(OTU %in% netprops_sperm.crossdomain.cottonvsoybean$hubs$hubs2)
soybean.bac.hubs
cotton.bac.hubs <- tax.bac %>%
  subset(OTU %in% netprops_sperm.crossdomain.cottonvsoybean$hubs$hubs1)
cotton.bac.hubs
soil.bac.hubs <- tax.bac %>%
  subset(OTU %in% netprops_sperm.crossdomain.cottonvsoil$hubs$hubs2)
soil.bac.hubs # no bacterial hubs in soil? 

# Fungi
soybean.fun.hubs <- tax.fun %>%
  subset(OTU %in% netprops_sperm.crossdomain.cottonvsoybean$hubs$hubs2)
soybean.fun.hubs
cotton.fun.hubs <- tax.fun %>%
  subset(OTU %in% netprops_sperm.crossdomain.cottonvsoybean$hubs$hubs1)
cotton.fun.hubs
soil.fun.hubs <- tax.fun %>%
  subset(OTU %in% netprops_sperm.crossdomain.cottonvsoil$hubs$hubs2)
soil.fun.hubs