---
title: "Transmission Analysis"
output: html_document
date: "2024-06-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#####Transmission Analysis#####
```{r}
###### Libraries #####
library(phyloseq)
library(vegan)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(microbiome)
library(ggVennDiagram)
library(VennDiagram)
library(dplyr)

##### Set global options #####

# no scientific notation
options(scipen=10000) 

# color blind pallet used throughout 
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
ibm.cbb <- c("#648FFF", "#785EF0", "#DC267F", "#FE6100", "#FFB000")
tol.cbb <- c("#332288", "#117733", "#44AA99", "#88CCEE", "#DDCC77", "#CC6677", "#AA4499", "#882255")
```
#Load in RDS#
```{r}
bac_sperm <- readRDS("2024-01-24_SpermosphereSteaming_Bacteria/RDS/Bacteria_spermosphere_nonnorm_2024-01-24.rds") ##Using Non-normalized Reads
bac_sperm_m <- data.frame(bac_sperm@sam_data)
```
#Transmission analysis #

#### The purpose of this analysis is to determine which OTUs are only present in the spermosphere samples compared to bulk soil. Which may indicate that they came from the seed.
#### We will transform the dataset to presence absence, then compared what was present in only spermospheres. 

## Bacteria ##
### Subsetting Samples ###
###Grouping of samples and Abundancies
```{r}
b <- bac_sperm %>%
  psmelt() %>% #combining the Physeq Object
  group_by(OTU, Trt, Time, Type) %>% #Grouping by different 
  summarize(sumabund = sum(Abundance)) #Has Zeros in It?
  
b$PA <- ifelse(b$sumabund > 0, 1, 0)
```
###Forming Presence absence based on Steamed Samples
```{r}
Soybean_Epiphyte.present <- b$OTU[b$PA == 1 & b$Trt == "Soybean Epiphytes"]
BulkSoil_Planting_Steamed.present <- b$OTU[b$PA == 1 & b$Trt == "Steamed " & b$Time == "Planting" & b$Type == "Bulk.Soil"]

Spermosphere_Planting_Steamed.present <- b$OTU[b$PA == 1 & b$Trt == "Steamed " & b$Time == "Planting" & b$Type == "Spermosphere"]
Spermosphere_17hrs_Steamed.present <- b$OTU[b$PA == 1 & b$Trt == "Steamed " & b$Time == "17hrs" & b$Type == "Spermosphere"]
Spermosphere_Rhizosphere_Steamed.present <- b$OTU[b$PA == 1 & b$Trt == "Steamed " & b$Time == "V2" & b$Type == "Rhizosphere"]


bS_Eonly <- setdiff(Soybean_Epiphyte.present, BulkSoil_Planting_Steamed.present)
b1 <- intersect(bS_Eonly, Spermosphere_Planting_Steamed.present)
b2 <- intersect(b1, Spermosphere_17hrs_Steamed.present)
b3 <- intersect(b2, Spermosphere_Rhizosphere_Steamed.present)
```
###Forming Presence absence based on Nonsteamed samples
```{r}
Soybean_Epiphyte.present <- b$OTU[b$PA == 1 & b$Trt == "Soybean Epiphytes"]
BulkSoil_Planting_NonSteamed.present <- b$OTU[b$PA == 1 & b$Trt == "Non.Steamed" & b$Time == "Planting" & b$Type == "Bulk.Soil"]

Spermosphere_Planting_NonSteamed.present <- b$OTU[b$PA == 1 & b$Trt == "Non.Steamed" & b$Time == "Planting" & b$Type == "Spermosphere"]
Spermosphere_17hrs_NonSteamed.present <- b$OTU[b$PA == 1 & b$Trt == "Non.Steamed" & b$Time == "17hrs" & b$Type == "Spermosphere"]
Spermosphere_Rhizosphere_NonSteamed.present <- b$OTU[b$PA == 1 & b$Trt == "Non.Steamed" & b$Time == "V2" & b$Type == "Rhizosphere"]

nbS_Eonly <- setdiff(Soybean_Epiphyte.present, BulkSoil_Planting_NonSteamed.present)
nb1 <- intersect(nbS_Eonly, Spermosphere_Planting_NonSteamed.present)
nb2 <- intersect(nb1, Spermosphere_17hrs_NonSteamed.present)
nb3 <- intersect(nb2, Spermosphere_Rhizosphere_NonSteamed.present)
```
###Correlating OTU to taxa
```{r}
nonsteamedtaxa <- bac_sperm@tax_table %>%
      data.frame() %>%
      subset(OTU %in% nb3)
steamedtaxa <- bac_sperm@tax_table %>%
      data.frame() %>%
      subset(OTU %in% b3)

```

# ggVennDiagram return a ggplot object, the fill/edge colors can be further modified with ggplot functions.
```{r}
###NonSteamed
x <- list(
  "Bulk Soil at Planting" = BulkSoil_Planting_NonSteamed.present, 
  "Seed Epiphytes" = Soybean_Epiphyte.present
  
)
ven <- ggVennDiagram(x, color = 1, lwd = 0.8, lty = 1) +
  #scale_fill_gradient(low="blue",high = "red") +
   +
  scale_x_continuous(expand = expansion(mult = .2)) +
  scale_fill_gradient(low = "#F4FAFE", high = "#4981BF") +
  coord_flip() +
  scale_color_manual(values = c("black", "black", "black"))
ven

###Steamed
y <- list(
   "Bulk Soil at Planting" = BulkSoil_Planting_Steamed.present,
  "Seed Epiphytes" = Soybean_Epiphyte.present
 
)
ven <- ggVennDiagram(y, color = 1, lwd = 0.8, lty = 1) +
  #scale_fill_gradient(low="blue",high = "red") +
  scale_x_continuous(expand = expansion(mult = .2)) +
  scale_fill_gradient(low = "#F4FAFE", high = "#4981BF") +
  coord_flip() +
  scale_color_manual(values = c("black", "black", "black"))
ven

```

##### Fungi ####
```{r}
fungi.filt <- core(fungi_sperm, detection = 0.01, prevalence = 0.1)
```
##Replace bac_Sperm with Fungi.filt

#### Steamed Spermosphere #####
```{r}
soybean.sub <- bac_sperm %>% 
  subset_samples(Crop == "Soybean")
otu.soybean <- soybean.sub@otu_table %>%
  as("matrix")
map.soybean <- soybean.sub@sam_data %>%
  as("data.frame")
otu_PA_soybean <- 1*((otu.soybean>0)==1)   
otu_occ_soybean <- data.frame(rowSums(otu_PA_soybean)); colnames(otu_occ_soybean) <- "Abundance"

otu_occ_soybean$OTU <- rownames(otu_occ_soybean)
otu_occ_soybean$Crop <- "Soybean"
```
##### Nonsteamed Spermospehre ##### 
#Future Logan (TWO PLANT SPECIES WERE DONE HERE MAKE IT ABOVE ###  INSTEAD)
```{r}
cotton.sub <- bac_sperm %>% 
  subset_samples(Crop == "Cotton ")
otu.cotton <- cotton.sub@otu_table %>%
  as("matrix")
map.cotton <- cotton.sub@sam_data %>%
  as("data.frame")
otu_PA_cotton <- 1*((otu.cotton>0)==1)   
otu_occ_cotton <- data.frame(rowSums(otu_PA_cotton)); colnames(otu_occ_cotton) <- "Abundance"

otu_occ_cotton$OTU <- rownames(otu_occ_cotton)
otu_occ_cotton$Crop <- "Cotton"
```
##### Bulk Soil Steamed #####
```{r}
bulk.sub <- bac_sperm %>% 
  subset_samples(Crop == "Bulk Soil")
otu.bulk <- bulk.sub@otu_table %>%
  as("matrix")
map.bulk <- bulk.sub@sam_data %>%
  as("data.frame")
otu_PA_bulk <- 1*((otu.bulk>0)==1)   
otu_occ_bulk <- data.frame(rowSums(otu_PA_bulk)); colnames(otu_occ_bulk) <- "Abundance"

otu_occ_bulk$OTU <- rownames(otu_occ_bulk)
otu_occ_bulk$Crop <- "Bulk Soil"
```
##### Bulk Soil Non_Steamed #####
```{r}
bulk.sub <- bac_sperm %>% 
  subset_samples(Crop == "Bulk Soil")
otu.bulk <- bulk.sub@otu_table %>%
  as("matrix")
map.bulk <- bulk.sub@sam_data %>%
  as("data.frame")
otu_PA_bulk <- 1*((otu.bulk>0)==1)   
otu_occ_bulk <- data.frame(rowSums(otu_PA_bulk)); colnames(otu_occ_bulk) <- "Abundance"

otu_occ_bulk$OTU <- rownames(otu_occ_bulk)
otu_occ_bulk$Crop <- "Bulk Soil"
```
##### Rhizosphere Non-Steamed #####
```{r}
bulk.sub <- bac_sperm %>% 
  subset_samples(Crop == "Bulk Soil")
otu.bulk <- bulk.sub@otu_table %>%
  as("matrix")
map.bulk <- bulk.sub@sam_data %>%
  as("data.frame")
otu_PA_bulk <- 1*((otu.bulk>0)==1)   
otu_occ_bulk <- data.frame(rowSums(otu_PA_bulk)); colnames(otu_occ_bulk) <- "Abundance"

otu_occ_bulk$OTU <- rownames(otu_occ_bulk)
otu_occ_bulk$Crop <- "Bulk Soil"
```
##### Rhizosphere Steamed #####
```{r}
bulk.sub <- bac_sperm %>% 
  subset_samples(Crop == "Bulk Soil")
otu.bulk <- bulk.sub@otu_table %>%
  as("matrix")
map.bulk <- bulk.sub@sam_data %>%
  as("data.frame")
otu_PA_bulk <- 1*((otu.bulk>0)==1)   
otu_occ_bulk <- data.frame(rowSums(otu_PA_bulk)); colnames(otu_occ_bulk) <- "Abundance"

otu_occ_bulk$OTU <- rownames(otu_occ_bulk)
otu_occ_bulk$Crop <- "Bulk Soil"
```
#### Combine ######
```{r}
combined.occ <- rbind.data.frame(otu_occ_bulk, otu_occ_cotton, otu_occ_soybean)
combined.occ$pres.abs <- ifelse(combined.occ$Abundance > 5, 1, 0)
combined.occ$otu.pres <- interaction(combined.occ$Crop, combined.occ$pres.abs)

BulkSoil.absent <- combined.occ$OTU[combined.occ$otu.pres == "Bulk Soil.0"]
Soybean.absent <- combined.occ$OTU[combined.occ$otu.pres == "Soybean.0"]
Cotton.absent <- combined.occ$OTU[combined.occ$otu.pres == "Cotton.0"]

BulkSoil.present <- combined.occ$OTU[combined.occ$otu.pres == "Bulk Soil.1"]
Soybean.present <- combined.occ$OTU[combined.occ$otu.pres == "Soybean.1"]
Cotton.present <- combined.occ$OTU[combined.occ$otu.pres == "Cotton.1"]

x <- list(
  "Cotton" = Cotton.present,
  "Soybean" = Soybean.present,
  "Bulk Soil" = BulkSoil.present
)
```
# ggVennDiagram return a ggplot object, the fill/edge colors can be further modified with ggplot functions.
```{r}
ven <- ggVennDiagram(x, color = 1, lwd = 0.8, lty = 1) +
  #scale_fill_gradient(low="blue",high = "red") +
  scale_x_continuous(expand = expansion(mult = .2)) +
  scale_fill_gradient(low = "#F4FAFE", high = "#4981BF") +
  scale_color_manual(values = c("black", "black", "black"))
ven
```
###Calculations and Such###
```{r}
overlap <- calculate.overlap(
  x = list(
    "Cotton" = Cotton.present,
    "Soybean" = Soybean.present,
    "Bulk Soil" = BulkSoil.present));

length(overlap[[1]]) # intersect of all
length(overlap[[2]]) # coton and soybean spermosphere 
length(overlap[[3]]) # bulk soil and cotton
length(overlap[[4]]) # soybean and bulk soil
length(overlap[[5]]) # Cotton unique
length(overlap[[6]]) # soybean unique
length(overlap[[7]]) # Bulk soil unique

soybean.unique <- fungi_sperm@tax_table %>%
  as.matrix() %>%
  as.data.frame() %>%
  subset(OTU %in% overlap[[5]]) %>%
  mutate(Crop = "Soybean")

cotton.unique <- fungi_sperm@tax_table %>%
  as.matrix() %>%
  as.data.frame() %>%
  subset(OTU %in% overlap[[6]]) %>%
  mutate(Crop = "Cotton")

bulk.unique <- fungi_sperm@tax_table %>%
  as.matrix() %>%
  as.data.frame() %>%
  subset(OTU %in% overlap[[7]]) %>%
  mutate(Crop = "Bulk Soil")

soy.cot <- fungi_sperm@tax_table %>%
  as.matrix() %>%
  as.data.frame() %>%
  subset(OTU %in% overlap[[2]]) %>%
  mutate(Crop = "Soybean and Cotton")

combined.unique <- rbind.data.frame(soy.cot, soybean.unique, cotton.unique, bulk.unique)

combined.unique %>%
  group_by(Crop, Phylum) %>%
  count() %>%
  mutate(Genus.label = ifelse(n >= 1, Phylum, "Other")) %>%
  ggplot(aes(x = Crop, y = n, fill = Genus.label)) +
  geom_bar(position="fill", stat = "identity") + 
  theme_classic() + 
  xlab("")+
  ylab("Proportion")+
  scale_fill_manual(values=c(cbbPalette, ibm.cbb, tol.cbb, "grey", "green"))
```


